---
title: "STAT 4410/8416 Homework 5"
author: "Danuwar Ramesh"
date: "Due on Nov 28, 2020"
output:
  pdf_document: 
    latex_engine: xelatex
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(fig.align='center', dpi=100, message=FALSE, warning=FALSE, cache=TRUE)
output <- opts_knit$get("rmarkdown.pandoc.to")
if(!is.null(output)) {
  if (output=="html") opts_chunk$set(out.width = '400px') else
    opts_chunk$set(out.width='.6\\linewidth')
}
```

1. **Working with databases:** Please follow the instruction below before answering the questions:

* Install the package `sqldf` using `install.packages('sqldf')`
* Import the library using `library('sqldf')`
* Read the file `https://raw.githubusercontent.com/dsindy/kaggle-titanic/master/data/train.csv ` and store it in `titanic`

```{r, echo=FALSE}
library(data.table)
titanic <- fread('https://raw.githubusercontent.com/dsindy/kaggle-titanic/master/data/train.csv')
```


We can now start writing SQL Script using `SQLDF library` right inside R. See example below:

```{r}
library(sqldf)

sqldf("SELECT passengerid, name, sex
      FROM titanic
      limit 5", drv="SQLite")
```


Answer the following questions. Write SQL Script where applicable.

a. What is a Database Management System? Name few types of Database Management System.
Answer: Database management system is the system that allows to store, modify, and extract information from a database.The different types of databasae management system are: mysql, oracle, ms sql, db2, ms access etc

b. Provide 3 verbs of SQL? Please write what they do
Answer:
Select:The SELECT statement is used to select data from a database.
Update:The SQL UPDATE Query is used to modify the existing records in a table
Delete:To remove one or more rows from a table completely, you use the DELETE statement
INSERT:used to add new rows of data to a table in the database

c. What does the following command do in MySQL?
    i. `show databases;
     Answer: To display the available databases the command is show.
    ii. `show tables;`
    Answer:It helps to list the tables for the current/specified database or schema or accross the entire account.
    
d. Write `SQL` script to answer the following questions based on titanic data. Display the results of your script.
    iii. What is the average age of passengers who survived? Group the data by Sex. Display only the column `Sex`, `AverageAge`
```{r}
kable(sqldf("SELECT Sex, 
                    avg(Age) as AverageAge
      FROM titanic
      WHERE Survived = 1
      group by Sex", drv="SQLite"))
```
    
    
    iv. What is the percentage of passengers who survived in each Passenger Class or `Pclass`? Group the data by `Sex`. Display Pclass, Sex, percentage value. 
```{r}
kable(sqldf("SELECT count(Survived) as total_rows_survived
      FROM titanic", 
            drv="SQLite"))
```
```{r}
sqldf("SELECT pclass, 
      sex, cast(sum(survived) as real) / count(survived) * 100 as avg_survived_percent
      FROM titanic
      GROUP BY pclass, sex;"
      , drv="SQLite")
```

    
    v. What is the average age of all the passenger (survived and not survived)? Group the data by `Pclass`, `Sex`, `Survived`. After that use `ggplot` to generate a line plot to show average age vs pclass, facet by sex and color it by survived.
```{r}
q1_iii <- sqldf("SELECT Pclass, 
                        Sex, 
                        Survived, 
                        avg(Age) as AverageAge
      FROM titanic
      group by Pclass, Sex, Survived", drv="SQLite")

kable(q1_iii)
```
```{r}
library(ggplot2)
ggplot(data=q1_iii, aes(x=AverageAge, y=Pclass, color=Survived, group=Survived))+
  geom_line()+
  geom_point()+
  facet_grid(~Sex)+
  xlab("Average Age")+
  ylab("Passenger Class")+
  labs(color='Survived')
```
    
    vi. What is the name, age, sex and pclass of the 5 oldest and 5 youngest persons who died?
```{r}
# 5 Youngest person who died
kable(sqldf(" SELECT Name, Age, Sex, Pclass
        FROM titanic
        WHERE Age IS NOT NULL OR age <= 0
        AND Survived = 0
        ORDER BY Age ASC
        LIMIT 5", drv="SQLite"))
```
    
```{r}
# 5 Oldest person who died
kable(sqldf(" SELECT name, age, sex, pclass
        FROM titanic
        WHERE age IS NOT NULL OR age <= 0
        AND Survived = 0
        ORDER BY age DESC
        LIMIT 5", drv="SQLite"))
```
    
    vii. On average which Passenger Class is more expensive?
```{r}
kable(sqldf(" SELECT Pclass, avg(Fare) as AverageFare
        FROM titanic
        GROUP BY Pclass", drv="SQLite"))
```


2. **Extracting twitter data;** In this problem we would like to extract data from twitter. For this refer to the documentation in the following link.

`https://github.com/geoffjentry/twitteR/`

a. **Twitter API set up** Set up twitter API using any of the following methods. Make sure you installed all the packages as mentioned in the class.

**Method 1:** Read Getting Started section of the above link and create a twitter application by going to the link `https://apps.twitter.com/`. Once you created your application connect twitter from `R` using the secrets and keys obtained from your twitter application. 


```{r, eval=FALSE}
library(twitteR)
api_key <- "your api key"  
api_secret <- "your api secret" 
access_token <- "your access tocken"
access_token_secret <- "your access tocken secret"

setup_twitter_oauth(api_key,api_secret,access_token,
                    access_token_secret)
```

**Method 2:** If you don't like creating an account with twitter and going through all the trouble, you can use my keys (ssh, don't tell anyone). For this download the `hw5-twitter-auth` file from blackboard and load it as follows.

```{r}
load("hw5-twitter-auth")
library(base64enc)
library(httr)
library(twitteR)
setup_twitter_oauth(api_key,api_secret,access_token,
                    access_token_secret)
```

b. Now search twitter messages for "data science job". Display few job informations.
```{r}
datascience_jobs_tweets <- twitteR::searchTwitter("data science job")
head(datascience_jobs_tweets)
```




c. Search 300 tweets using the hash tag `#chess` and save them in an object called `rTweets`. Show the top 7 sources of tweets (such as android or iphone) in a ordered bar plot. 
```{r}
# Search 300 tweets with (#chees)
rTweets <- twitteR::searchTwitter("#chess", n=300)

# Get the source from where the tweet came from
tweet_sources_raw <- sapply(rTweets, function(x) x$getStatusSource())

# Remove html tags that is on the sources and only get the value between the tags
# This is the source
library(stringr)
regex_pattern <- '[^">]+\\</a>'
tweet_sources_text <- gsub("</a>","", unlist(str_extract(tweet_sources_raw, regex_pattern)))

# Display few sources
head(tweet_sources_text)

# We now have the source from which the 300 tweets were being sent
# Now lets get the frequency count of the sources of distict 300 tweets 
tweet_sources_text_FreqDist <- table(tweet_sources_text)

# Display few tweets with Frequenct Distribution
head(tweet_sources_text_FreqDist)

``` 
```{r}
library(ggplot2)
# Convert the Frequency Distirbution data into ta nice Data Frame
# Also sort the frequency data in decreasing order at the same time
library(dplyr)
rTweetsDF_sources <- data.frame(tweet_sources_text_FreqDist) %>% arrange(desc(Freq))

# Display few data
kable(head(rTweetsDF_sources))

# We only want the top 7 data
# Now lets subset the data
rTweetsDF_sources_top7 <- head(rTweetsDF_sources, 7)

# Display top 7 data
rTweetsDF_sources_top7


# Use ggplot to plot the ordered bar plot
ggplot(rTweetsDF_sources_top7, aes(x=reorder(tweet_sources_text, Freq), y =Freq ))+
  geom_bar(stat="identity")+
  xlab("Sources of Tweet")+
  ylab("Total Frequency")+
  coord_flip() 
```  
d. Notice that the object `rTweets` is a list. Convert it into a data frame using function `twListToDF` and store it in an object called `dTweets`. Display some data from `dTweets`.
```{r}
dTweets <- twListToDF(rTweets)
head(dTweets)
```


e. `dTweets` has a column showing the time the tweet was created. Generate a plot showing number of tweets on each of the hours. Add a smooth line overlaid on your plot.
```{r}
library(lubridate)

# convert the time into the format ymd_hms()
hour <- data.frame(hour(ymd_hms(dTweets$created)))

# add a new column in the dTweet with hour value
dTweets <- data.frame(hour, dTweets)

# rename the newly added hour column
colnames(dTweets)[1] <-"hourValue"

# filter NA values if any
dTweetsHourCount <- dTweets %>%
  filter(!is.na(hourValue)) %>%
  group_by(hourValue) %>%
  tally()

# Display Tweets per hour table
kable(dTweetsHourCount)

# Generate a plot showing number of tweets on each of the hours
# Add a smooth line overlaid on your plot
ggplot(dTweetsHourCount, aes(x=hourValue, y=n))+
  geom_smooth(method = 'loess')+
  geom_point()+
  xlab("Hour Value")+
  ylab("Total tweets per hour")
``` 


f. Arrange the dataframe `dTweets` based on the `retweetCount`. While doing this select only columns `text, screenName, retweetCount`. Store the data in a object called `mostTweets`. Display five texts that are most retweeted.
```{r}
#Arrange the dataframe `dTweets` based on the `retweetCount`
# While doing this select only columns `text, screenName, retweetCount`
mostTweets <- dTweets %>%
  select(text, screenName, retweetCount) %>%
  arrange(desc(retweetCount))

head(mostTweets$text,5)
```


g. Generate a bar chart showing top 15 screen names and count of retweets from `mostTweets`. Order the bars based on the retweet counts.
```{r}
top15_mostTweets <- head(mostTweets, 15)

ggplot(top15_mostTweets, aes(x = reorder(screenName, retweetCount), y = retweetCount)) +
  geom_bar(stat= "Identity") +
  xlab("Twitter Screen Name") +
  ylab("Total retweet count") +
  coord_flip() +
  theme(plot.title = element_text(hjust = 0.5))
```

